Supporting Information:\\ Headgroup structure and cation binding in phosphatidylserine lipid bilayers

### 2019-08-20 Manuscript/manuscriptPS.tex
Updates related to the simulations with NBfix2 paraeters

POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13,kim16}  & 0.26  & 250 & 11190  & 53  & 298  & 200  & 180  & \citenum{POPC5POPS1withCaClCHARMM} \tabularnewline
POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13,kim16}  & 1.06  & 250 & 11174  & 214  & 298  & 200  & 180  & \citenum{POPC5POPS1with1MCaClCHARMM} \tabularnewline

>>>>

POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 0.26  & 250 & 11190  & 53  & 298  & 200  & 180  & \citenum{POPC5POPS1withCaClCHARMM} \tabularnewline
POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 1.06  & 250 & 11174  & 214  & 298  & 200  & 180  & \citenum{POPC5POPS1with1MCaClCHARMM} \tabularnewline
POPC:POPS (5:1)/Na$^{+}$  & CHARMM36 \cite{klauda10,venable13}  & 0  & 250  & 11207  & 0  & 320  & 400  & 300  & \citenum{POPC5POPS1CHARMMwithNBfixHan} \tabularnewline
POPC:POPS (5:1)/Na$^{+}$  & CHARMM36 \cite{klauda10,venable13} / NBfix2 \cite{han2018graph}  & 0.14  & 250 & 11190  & 28  & 320  & 440  & 300  & \citenum{POPC5POPS1CHARMMwithNBfixHan} \tabularnewline
POPC:POPS (5:1)/Na$^{+}$  & CHARMM36 \cite{klauda10,venable13} / NBfix2 \cite{han2018graph}  & 0.94  & 250 & 11174  & 189  & 320  & 440  & 300  & \citenum{POPC5POPS1CHARMMwithNBfixHan} \tabularnewline

### 2019-08-20 Manuscript/manuscriptPSsuppl.tex
Updates related to the simulations with NBfix2 paraeters

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} 
POPC:POPS (5:1) mixtures containing total 300 (250:50) lipids
%were simulated with the additional calcium concentrations.
%The data is available from Refs. \citenum{POPC5POPS1withCaClCHARMM} and \citenum{POPC5POPS1with1MCaClCHARMM}.
and 0.26~M and 1~M CaCl$_2$ were prepared using the CHARMM-GUI \cite{lee16,jo18}.
Systems were simulated using Gromacs 5 \cite{abraham2015gromacs} and CHARMM36 force field with the simulation parameters given
by the CHARMM-GUI at 298~K.

@@

\section{Calcium binding to POPC in CHARMM36 simulation with NBfix}\label{CHARMMcalciumNBfix}

The response of POPC headgroup order parameters to the CaCl$_2$ concentration
are underestimated in simulations of POPC:POPS (5:1) mixture with CHARMM36 when employing NBfix (as obtained from CHARMM-GUI in January 2018) for interactions between calcium and lipid oxygens \cite{kim16}
(Fig.~\ref{changesWITHCaClPS} in the main text),
indicating that the calcium binding to the bilayer is too weak with these
parameters. The response of headgroup order
parameters (Fig.~\ref{OP_CHARMM_CaCl_POPC_NBFix}) and the binding affinity
(Fig.~\ref{density_profile_CHARMM_CaCl_POPC_NBFix}) of calcium also to a pure
POPC bilayer are underestimated when using NBfix ions.  Notably, CHARMM36 simulations with the NBfix
terms \cite{venable13,kim16} predict similar binding affinity for sodium and calcium. Without the NBfix term, the calcium binding affinity to pure POPC lipid bilayers was overestimated in the CHARMM36 model \cite{catte16}.

%HA: this just repeats the above? 
%In conclusion, the special NBfix~\cite{kim16} for calcium, 
%incorporated in the parameters give by the CHARMM-GUI at the time
%of running the simulations (January 2018), leads to the underestimation of
%calcium binding affinity to pure PC and mixed PC:PS bilayers.

@@

\noindent
{\it Josef Melcr}
prepared tools for calculating order parameters, performed Lipid14/17 simulations with Gromacs MD simulation package,
and provided comments on the manuscript.

@@

\noindent
{\it Ricky Nencini}
ran the simulations of POPC systems using CHARMM36 with and without the NB-Fix corrections,
and analyzed the difference between bulk and buffer concentrations.

>>>>

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium using NBfix1 parameters.} 
POPC:POPS (5:1) mixtures containing total 300 (250:50) lipids
%were simulated with the additional calcium concentrations.
%The data is available from Refs. \citenum{POPC5POPS1withCaClCHARMM} and \citenum{POPC5POPS1with1MCaClCHARMM}.
and 0.26~M and 1~M CaCl$_2$ were prepared using the CHARMM-GUI in Janyary 2018 \cite{lee16,jo18}.
The ion parameters generated by CHARMM-GUI at the time are labeled here as NBfix1~\cite{kim16}.
Systems were simulated using Gromacs 5 \cite{abraham2015gromacs} and CHARMM36 force field with the simulation parameters given
by the CHARMM-GUI at 298~K. \\

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium using NBfix1 parameters.}
\todo{Ricky, please update the simulation methods.}
Systems were prepared using the CHARMM-GUI in August 2019 \cite{lee16,jo18}.
The ion parameters generated by CHARMM-GUI at the time are labeled here as NBfix2~\cite{han2018graph}.

@@

\clearpage
\section{Calcium binding to POPC in CHARMM36 simulation with NBfix1 parameters}\label{CHARMMcalciumNBfix}

The response of POPC headgroup order parameters to the CaCl$_2$ concentration
are underestimated in simulations of POPC:POPS (5:1) mixture with CHARMM36 when employing the
NBfix %(as obtained from CHARMM-GUI in January 2018)
parameters for the interactions between calcium and lipid oxygens \cite{kim16,han2018graph}
(Fig.~\ref{changesWITHCaClPS} in the main text), indicating that the calcium binding to the
bilayer is too weak with these parameters. The response of headgroup order
parameters (Fig.~\ref{OP_CHARMM_CaCl_POPC_NBFix}) and the binding affinity
(Fig.~\ref{density_profile_CHARMM_CaCl_POPC_NBFix}) of calcium
were underestimated also in simulations of pure POPC bilayer with the NBfix1 parameters~\cite{kim16}.
Notably, CHARMM36 simulations with the NBfix1 terms \cite{venable13,kim16} predict similar binding affinity for sodium and calcium.
Because the results for POPC:POPS (5:1) mixture were very similar with both NBfix parameters~\cite{kim16,han2018graph},
the NBfix2 parameters were not separately tested with pure POPC bilayers.
Without the NBfix parameters, the calcium binding affinity to pure POPC lipid bilayers was overestimated in the CHARMM36 model \cite{catte16}.

@@

\noindent
{\it Josef Melcr}
prepared tools for calculating order parameters, performed Lipid14/17 simulations with Gromacs MD simulation package,
prepared the TOC figure and provided comments on the manuscript.

@@

\noindent
{\it Ricky Nencini}
ran the simulations of POPC systems using CHARMM36 with and without the NB-Fix corrections,
ran the simulations of POPC:POPS (5:1) systems using CHARMM36 with the NBfix2 correction,
and analyzed the difference between bulk and buffer concentrations.
